---
title: "Intelligence and Education"
execute: 
  echo: false
  warning: false
---

Does the relationship between intelligence and education change over time? The following data was pulled from the General Social Survey using the gssr package. Mapping data was then used to visualize the relationship between intelligence (10-item vocabulary test called "Wordsum" was used as proxy for intelligence) and education level from 1974 to 2020. The strength of the relationship between intelligence and education seems to be decreasing with time.

```{r}
#| message: false
library(tidyverse)
library(gssr)
data(gss_all)
```

```{r}
# Prepare the data
gss_data <- gss_all %>%
  select(year, wordsum, educ) %>%
  filter(!is.na(wordsum) & !is.na(educ)) %>%
  mutate(decade = case_when(
    year >= 1970 & year < 1980 ~ "1970s",
    year >= 1980 & year < 1990 ~ "1980s",
    year >= 1990 & year < 2000 ~ "1990s",
    year >= 2000 & year < 2010 ~ "2000s",
    year >= 2010 & year < 2020 ~ "2010s",
    year >= 2020 ~ "2020s"
  ))
```

```{r}
# Create a scatter plot with a smooth line to show the relationship by decade
ggplot(gss_data, aes(x = educ, y = wordsum, color = decade)) +
  geom_jitter(alpha = 0.1) +  # Make points more transparent
  geom_smooth(method = "lm", se = FALSE, size = 1.5) +  # Increase line width
  labs(title = "Relationship between Wordsum and Education Level by Decade",
       subtitle = "The strength of the relationship seems to be decreasing with time",
       x = "Education Level (Years)",
       y = "Wordsum Score",
       color = "Decade") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
